Bayesian exponential random graph modeling of KWL Burial network

### 2020-12-03 analysis/paper/paper.Rmd
work on discussion section and run some more models

<!-- archaeological specific case study -->
In this study, we use a pericolonial model to explore the changes in social inequality at Kiwulan, an Iron Age site in northeastern Taiwan, which covers the time before the European arrival, the presence of the Europeans, including the Spanish and the Dutch, in the 17th century, and finally the Chinese in the 19th century [@Chen2007]. We assume that the social changes at Kiwulan would be supported by chronological differences in the structure of burial networks. We examine burial networks before and after the foreign contacts, and test our hypothesis that increased social inequality can be observed in the network after the European arrival. We ask: (1) does the observed burial data after the start of European presence resemble the of network our hypothesized model? (2) did European colonial activities in 17th century Taiwan result in the emergence of social inequality in an Indigenous society in ways that can be detected by burials networks? (3) what are the major variables affecting or forming the higher degree of popularity (e.g. a few nodes have more relationships with other nodes) in the European contact period that might hint at social heterogeneity during this period? By answering these questions, this study helps to expand the use of burials in understanding the indirect effects of a colonial presence on Indigenous groups. This is important to understand the reactions of Kiwulan residents in the context of indirect impacts of European colonization.

@@

<!-- comparison between two networks using patchwork bootstrap --> 
One key difference between the pre-European and post-European networks is their size, with x burials compared to y burials. To understand if this network size different affects the posterior distributions, we conducted a sensitivity test using vertex bootstrap inference on the networks. This allows a comparison for multiple networks of different sizes [@Chen2019]. Vertex bootstrap is a non-parametric method that conducts resampling for all vertices (i.e. node) of a network to quantify standard errors in the network statistics of interest. This enables the evaluation of uncertainty for networks and the comparison between networks by examining their confidence interval for the network population. We set the resampling size to 1,000 networks to compute network statistics, including density, clustering, and degree, for two networks. The result shows that the 95% confidence interval of each network statistic contains the observed network of both periods, which indicates our networks are relatively robust to differences in sample size. This suggests a very small sample size effect on both models. In addition, since the prior setting will have impacts on the posterior, we compared the posterior estimates with the same prior for all parameters, and those with the informative prior based on our assumptions for some parameters. In general, there are no significant changes in the posterior estimates using different sets of priors. But we can see the informative prior contributed to a better convergence of the MCMC according to the diagnostic plots showing more stationary distributions and decreased autocorrelation. <!-- re-write wither to justify why vertex bootstrap is the right method to answer this question, or use patchwork bootstrap and justify that-->

@@

<!-- briefly discuss Chinese burials -->
Age and gender attributes serve as indicators for distinguishing ascribed or achieved status in a society [citation].

>>>>

<!-- archaeological specific case study -->
In this study, we use a pericolonial model to explore the changes in social inequality at Kiwulan, an Iron Age site in northeastern Taiwan, which covers the time before the European arrival, the presence of the Europeans, including the Spanish and the Dutch, in the 17th century, and finally the Chinese in the 19th century [@Chen2007]. We assume that the social changes at Kiwulan would be supported by chronological differences in the structure of burial networks. We examine burial networks before and after the foreign contacts, and test our hypothesis that increased social inequality can be observed in the network after the European arrival. We ask: (1) does the observed burial data after the start of European presence resemble the network of our hypothesized model? (2) did European colonial activities in 17th century Taiwan result in the emergence of social inequality in an Indigenous society in ways that can be detected by burials networks? (3) what are the major variables affecting or forming the higher degree of popularity (e.g. a few nodes have more relationships with other nodes) in the European contact period that might hint at social heterogeneity during this period? By answering these questions, this study helps to expand the use of burials in understanding the indirect effects of a colonial presence on Indigenous groups. This is important to understand the reactions of Kiwulan residents in the context of indirect impacts of European colonization.

@@

<!-- comparison between two networks using patchwork bootstrap --> 
One key difference between the pre-European and post-European networks is their size, with `r pre_E` burials compared to `r post_E` burials. To understand if this network size different affects the posterior distributions, we conducted a sensitivity test using vertex bootstrap inference on the networks. This allows a comparison for multiple networks of different sizes [@Chen2019]. Vertex bootstrap is a non-parametric method that conducts resampling for all vertices (i.e. node) of a network to quantify standard errors in the network statistics of interest. This enables the evaluation of uncertainty for networks and the comparison between networks by examining their confidence interval for the network population. We set the resampling size to 1,000 networks to compute network statistics, including density, clustering, and degree, for two networks. The result shows that the 95% confidence interval of each network statistic contains the observed network of both periods, which indicates our networks are relatively robust to differences in sample size. This suggests a very small sample size effect on both models. In addition, since the prior setting will have impacts on the posterior, we compared the posterior estimates with the same prior for all parameters, and those with the informative prior based on our assumptions for some parameters. In general, there are no significant changes in the posterior estimates using different sets of priors. But we can see the informative prior contributed to a better convergence of the MCMC according to the diagnostic plots showing more stationary distributions and decreased autocorrelation. <!-- re-write wither to justify why vertex bootstrap is the right method to answer this question, or use patchwork bootstrap and justify that-->

@@

<!-- archaeological implication based on the results-->
The notable finding in our results is the change in network properties of Kiwulan burials from a more cohesive network (represented by higher transitivity and negative popularity) toward a more centralized network (lower transitivity and positive popularity) after the European presence during the 17th century. This confirms out hypothesized models of chronological differences in the structure of burials based on grave goods and burial attributes associated with demography and mortuary practices. The differences are in line with the theoretical assumption derived from the corporate-network continuum for explaining horizontal hierarchy that lacks characteristics of the traditional definition of vertical hierarchy but shows some level of social inequality [@Feinman2000]. This theoretical framework works better to understand the small scale societies and the strategies for gaining power between ambitious individuals in the pericolonical context.

<!-- methodological implication-->
The ERGM network analysis within a Bayesian framework provides a robust way to examine relations between burials based on grave goods and attributes associated with demography and mortuary practices. The adoption of Bayesian inference on ERGM can solve the issues of computational tractability and degeneracy in parameter estimation in ERGM [@Caimo2017]. In addition, a Bayesian framework reduced the effects of small sample size or missing data commonly present in archaeological data by incorporating prior information and MCMC estimations. We demonstrated a plausible and robust method to test the hypothesized model of chronological changes in burial structures with probabilistic interpretation. It is important to note that although some network statistics of nodal attributes do not highly match our assumption, the Bayesian approach is still an efficient and effective method for evaluating the endogenous network effects. This closely related to the overall structure of burials that can be viewed as a representation of a given society.

<!-- more discussion of nodal attributes -->
Age and gender attributes serve as indicators for distinguishing ascribed or achieved status in a society [citation].
